---
title: "Karlbom"
author: "John Moshtaghi-Svensson"
output: 
    html_document:
        highlight: pygments
        theme: sandstone
        css: style.css
        toc: true
        code_folding: show
editor_options: 
  chunk_output_type: inline
---

__Datum__: `r lubridate::today()`

```{r setup, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
library(dplyr)
library(stringr)
library(glue)
library(janitor)
library(here)
library(scales)
library(ggplot2)
library(lubridate)
library(rlang)
library(plotly)
library(tidyr)

source(".Rprofile")
source("crcbase_utils/helpers.R")
source("crcbase_utils/crud/crcbase_read_and_write.R")
source("crcbase_utils/crud/setup_project_crud.R")
source("crcbase_utils/process/fix_scrcr_dubletter.R")
source("crcbase_utils/process/process_scrcr.R")
source("crcbase_utils/process/process_lisa.R")
# Loads a list called definitions containtin procedure_codes
source("vars/definitions.R")


test_for_k()
theme_set(theme_bw())

knitr::opts_chunk$set(
    # root.dir = '~/R/project',
    warning = FALSE,
    message = FALSE,
    warning = FALSE,
    echo = TRUE,
    include = TRUE
)

if (!exists("params")) {
    params <- list(
        data_date = "20220207", # when was the original data created
        project_name = project_name # from .Renviron, loaded in setup_project_crud.R
    )
}
```

# Find exposed
Apply inclusion and exlusion criteria for the exposed. Identified from SCRCR

```{r find exposed}
# Find the exposed
scrcr_clean <- read_only_from_crcbase_derived_csv("scrcr_clean.csv", sep = ";", dec = ".") # TODO change to , when updating data
scrcr <- process_scrcr(scrcr_clean)
```

### Inclusion criteria

#### 0. Include first ever crcr only
Includes synchronous tumours. Handled later
Previous crc identified from ipr i handled in `2_create_analysisdata.rmd`

```{r first ever}
scrcr[, first_diagdate_scrcr := min(diagdate_scrcr, na.rm = TRUE), by = "lopnr"]
nrow(scrcr)
scrcr_first <- scrcr[diagdate_scrcr == first_diagdate_scrcr]
nrow(scrcr_first)
```

#### 1. proceduredate 2007:2016

```{r inclusion study period diagdate}
tabyl(year(scrcr$proceduredate))
scrcr_first <- scrcr_first[year(proceduredate) %in% 2007:2016]
nrow(scrcr_first)
```

#### 2. diagage >= 18
As we dont have birthdate available, I am not able to calculate age at proceduredate. Only age at diagdate_scrcr is available.

```{r inclusion diagage}
scrcr_first %>%
    ggplot(aes(x = diagage)) +
    geom_histogram(binwidth = 1) +
    geom_vline(xintercept = 18, linetype = "dashed", color = "red") +
    ggtitle("Age distribution for exposed")
```

#### 3. curative_proc == 1 \
(a2_kurativ_v_rde, curative procedure)

```{r inclusion curative proc}
tabyl(scrcr_first, curative_proc)
scrcr_first <- scrcr_first[curative_proc == 1]
nrow(scrcr_first)
```


#### 4. stage in 1-3 or cm == 0 
(cm == cM stage). Stage derived from cm, ct, cp

```{r inclusion stage}
tabyl(scrcr_first, stage)
scrcr_first <- scrcr_first[stage %in% 1:3]
nrow(scrcr_first)
```

#### 5. Bowel resection 
(procedure_type eq 3)

```{r inclusion procedure_type}
tabyl(scrcr_first, procedure_type)
scrcr_first <- scrcr_first[procedure_type == 3]
nrow(scrcr_first)
```

#### 6. Remove synchronous tumours

Keep the row with the most information

```{r remove synchronous tumours}
scrcr_first <- fix_scrcr_dubletter(scrcr_first)
```

```{r}
write_to_project_original(scrcr_first, "uk_scrcr_20220207.csv")
```

## Studypopulation
Add comparators

```{r add comparators}

studypopulation <- read_studypopulation(exp_lopnrs = scrcr_first$lopnr)
if (nrow(studypopulation)) {
    warning("studypopulation has 0 rows. Check read")
}
# ADD proceduredate from scrcr
studypopulation <- left_join(
    studypopulation,
    scrcr_first[, .(lopnr, diagdate_scrcr, proceduredate)],
    by = c("strata" = "lopnr", "match_date" = "diagdate_scrcr")
)

studypopulation[, index_date := proceduredate]
nrow(studypopulation)

write_to_project_original(studypopulation, "studypopulation.csv")
```

# Datasets

## SBO Surgeries

```{r sbo surgeries}
# SBO surgeries directly from ipr (not sed) ----
# all sbo surgeries. Not just the ones before. dont have index date available here.
ipr_clean <- read_only_from_crcbase_derived_csv("ipr_clean.csv", sep = ",", dec = ".", lopnr = studypopulation$lopnr)
ipr <- ipr_clean %>% select(lopnr, indate, hdia, dia1:dia30, surgerycodes)

# find the relevant codes (op)
# patterns for operations. Should allow optional leading whitespace
sbo_op_pattern1 <- definitions$procedure_codes[["sbo_operation_alt1"]] %>% str_c(collapse = "| ?")
sbo_op_pattern2 <- definitions$procedure_codes[["sbo_operation_alt2"]] %>% str_c(collapse = "| ?")

ipr[, `:=`(
    sbo_op1 = str_detect(surgerycodes, sbo_op_pattern1),
    sbo_op2 = str_detect(surgerycodes, sbo_op_pattern2)
)]
ipr[, .(sum(sbo_op1), sum(sbo_op2))]
sbo_surgeries <- ipr[sbo_op1 == 1 | sbo_op2 == 2]
nrow(sbo_surgeries)


# Find rows with ileus icd code
ileus_pattern <- definitions$icd_codes$ileus %>% str_c(collapse = "|")

# looks through all the dia columns and returns a vector with one value per row indicating if any ileus icds are present
sbo_surgeries$ileus_dia <- find_pattern_in_columns(select(sbo_surgeries, contains("dia")), ileus_pattern)
sbo_surgeries_dias <- sbo_surgeries[ileus_dia == TRUE]
nrow(sbo_surgeries_dias)
sbo_surgeries_dias[, .(
    nrow = .N,
    sum(sbo_op1),
    sum(sbo_op2),
    sum(sbo_op1 == 1 | sbo_op1 == 1)
)]
```

## Migrations
```{r}
# type 1 = immigration
# type 2 = emmmigration
migrations <- read_only_from_crcbase_derived_csv("migrations_clean.csv", lopnrs = studypopulation$lopnr, sep = ",", dec = ".") # TODO change seperator to ";"

migrations[, migration_type := ifelse(type == 2, "emmigration", "imigration")]
migrations <- left_merge(migrations, studypopulation[, .(lopnr, index_date)], by = "lopnr")
migrations <- migrations[migrationdate >= index_date]
write_to_project_original(migrations, "uk_migrations_20220207.csv")
```