---
title: "Karlbom"
author: "John Moshtaghi-Svensson"
output: 
    html_document:
        highlight: pygments
        theme: sandstone
        css: style.css
        toc: true
        code_folding: show
editor_options: 
  chunk_output_type: inline
---

__Datum__: `r lubridate::today()`

```{r setup, include=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
library(data.table)
library(dplyr)
library(stringr)
library(glue)
library(janitor)
library(here)
library(scales)
library(ggplot2)
library(lubridate)
library(rlang)
library(plotly)

source(".Rprofile")
source("crcbase_utils/helpers.R")
source("crcbase_utils/crud/setup_project_crud.R")
source("crcbase_utils/crud/read_from_variable_lists.R")
source("crcbase_utils/process/fix_scrcr_dubletter.R")
source("crcbase_utils/process/process_scrcr.R")
source("crcbase_utils/process/process_lisa.R")
# Loads a list called definitions containtin procedure_codes
source("vars/definitions.R")

test_for_k()

theme_set(theme_bw())

knitr::opts_chunk$set(
    # root.dir = '~/R/project',
    warning = FALSE,
    message = FALSE,
    warning = FALSE,
    echo = TRUE,
    include = TRUE
)

if (!exists("params")) {
    params <- list(
        end_of_study = "2016-12-31",
        data_date = "20220207", # when was the original data created
        project_name = project_name # from .Renviron, loaded in setup_project_crud.R
    )
}
```

# Description
Combines all data tables in project/data/original into one dataset and create all necessary new variables

# 1. Datasets

Data was created in sas program `uk_20210609.sas` under project/programs of r script `1_create_original_data.R` found in this project folder.

```{r read all data}
# Charlson for population
cci_exposed_raw <- read_from_project_original(glue("uk_cci_{params$data_date}.csv"))
cci_comparators_raw <- read_from_project_original(glue("uk_cci_comparators_{params$data_date}.csv"))

# Deaths for population
cdr_raw <- read_from_project_original(glue("uk_cdr_{params$data_date}.csv"))

# all iipr rows (before and after diagdate) indicating ileus, pain or hernia
# - ileus ("K565" "K566" "K567")
# - ab_pain (R10.0, R10.3, R10.4)
# - ab_hernia (K40.0, K40.3, K41.0, K41.3, K42.0, K43.0, K43.3, K43.6, K44.0, K45.0, K46.0)
ileus_raw <- read_from_project_original(glue("uk_ileus_{params$data_date}.csv"))

# all migrations for poulation
migrations_raw <- read_from_project_original(glue("uk_migrations_{params$data_date}.csv"))

# scrcr for population
scrcr_raw <- read_from_project_original(glue("uk_scrcr_{params$data_date}.csv"))

# all ipr row with ip codes specified in study plan (lots of codes, )
# Used to identify:
#- stoma closure IPR ONLY (op=JFG00, JFG10, JFG20, JFG23, JFG29, JFG26,JFG30, JFG33, JFG36)
surgeries_raw <- read_from_project_original(glue("uk_surgeries_{params$data_date}.csv")) # TODO still 0 rows

# Previous colorectal cancer (c18-c20) used for exclusion
previous_cr_cancer_raw <- read_from_project_original(glue("uk_previous_crc_{params$data_date}.csv"))
previous_cr_cancer <- previous_cr_cancer_raw[diagdate_scr < diagdate_scrcr]

# previous abdominal surgeries
previous_abd_surgeries_raw <- read_from_project_original(glue("uk_previous_abd_surgeries_{params$data_date}.csv"))

# first occurrence of ("C784" "C785" "C786" "C787") in patreg after diagdate_scrcr
# used for cencoring, i think
relapse_raw <- read_from_project_original(glue("uk_relapse_pr_{params$data_date}.csv")) # TODO change to .csv

# Abdominal hernia and abdominal pain from IPR
abd_hernia_raw <- read_from_project_original(glue("uk_abd_hernia_{params$data_date}.csv"))

# Lisa for educational level
lisa_raw <- read_from_project_original(glue("uk_lisa_{params$data_date}.csv"))

# Study population
studypop <- read_from_project_original("studypopulation.csv") # Created from studypopulation_comparators.sas7bdat and scrcr after inclusion and exlusion criteria
```

```{r import ables description}
import_tables <- list(
    cci_exposed_raw = cci_exposed_raw,
    cci_comparators_raw = cci_comparators_raw,
    cdr_raw = cdr_raw,
    ileus_raw = ileus_raw,
    migrations_raw = migrations_raw,
    scrcr_raw = scrcr_raw,
    surgeries_raw = surgeries_raw,
    previous_cr_cancer = previous_cr_cancer,
    previous_abd_surgeries_raw = previous_abd_surgeries_raw,
    relapse_raw = relapse_raw,
    abd_hernia_raw = abd_hernia_raw,
    studypop = studypop
)

exp_lopnr <- studypop[case == "exposed"]$lopnr
comp_lopnr <- studypop[case == "unexposed"]$lopnr
purrr::imap_dfr(import_tables, function(tbl, name) {
    tibble::tibble(
        name = name,
        nrow = nrow(tbl),
        ncol = ncol(tbl),
        n_unique_lopnr = length(unique(tbl$lopnr)),
        pop_in = scales::percent(mean(unique(studypop$lopnr) %in% tbl$lopnr)),
        exp_in = scales::percent(mean(unique(exp_lopnr) %in% tbl$lopnr)),
        comp_in = scales::percent(mean(unique(comp_lopnr) %in% tbl$lopnr))
    )
})
```

# 2. STUDY POPULATION

Most inclusion/exclusion criterion are alrady applied in the sas-file.  Will make controls here.

Applies in this file:
- previous colorectal cancer


```{r exposed popultion}
scrcr_processed <- process_scrcr(scrcr_raw)
scrcr_first <- fix_scrcr_dubletter(scrcr_processed)
```

### Inclusion criteria

__scrcr_1to1__ data in the sas program

#### 1. diagdate_scrcr 2007:2016
```{r inclusion study period diagdate}
scrcr_first %>% tabyl(diagyear_scrcr)
scrcr_first <- scrcr_first[year(diagdate_scrcr) %in% 2007:2016]
```

#### 2. diagage >= 18
```{r inclusion diagage}
scrcr_first %>%
    ggplot(aes(x = diagage)) +
    geom_histogram(binwidth = 1) +
    geom_vline(xintercept = 18, linetype = "dashed", color = "red") +
    ggtitle("Age distribution for exposed")
```

#### 3. curative_proc == 1 \
(a2_kurativ_v_rde, curative procedure)

```{r inclusion curative proc}
tabyl(scrcr_first, curative_proc)
scrcr_first <- scrcr_first[curative_proc == 1]
```


#### 4. stage in 1-3 or cm == 0 
(cm == cM stage). Stage derived from cm, ct, cp

```{r inclusion stage}
tabyl(scrcr_first, stage)
scrcr_first <- scrcr_first[stage %in% 1:3]
```

#### 5. Bowel resection 
(procedure_type eq 3)

```{r inclusion procedure_type}
tabyl(scrcr_first, procedure_type)
scrcr_first <- scrcr_first[procedure_type == 3]
```

### Exclusion criteria

#### 1. Previous Colorectal cancer

Identified from the swedish cancer register. Any before `diagdate_scrcr` plus 3months _a bit arbitrary but almost all have a cancer diagnosis before diagdate_scrcr. Im guessing it is  the same one. Need a washout period_
codes C18, C19, C20

- This is applied to the whole population (exposed and comparators)

```{r previous colorectal cancer}

previous_cr_cancer[, diagdate_scr_3m := diagdate_scr %m+% months(3)]
previous_cr_cancer[, timediff_m := lubridate::time_length(as_date(diagdate_scrcr) - as_date(diagdate_scr), unit = "months")]
prev_crc <- previous_cr_cancer[timediff_m > 3, .(lopnr, diagdate_scr, diagdate_scrcr, timediff_m)]
prev_crc_first <- prev_crc[, .(diagdate_crc = min(diagdate_scr), timediff_m = min(timediff_m)), by = c("lopnr", "diagdate_scrcr")]

#  plot distribution
qplot(prev_crc_first$difftime_m) + ggtitle("months since previous colorectal cancer")

# add to exposed statsu
prev_crc_first <- merge(prev_crc_first, studypop, by = c("lopnr", "diagdate_scrcr"), all.x = TRUE, all.y = FALSE)
# n exluded
prev_crc_first[, .N, by = "case"]

studypop <- studypop[!lopnr %in% prev_crc_first$lopnr]
```


### Exposed

Cancer surgery (bowel resection[procedure_type==3]) after colorectal cancer. Identified from SCRCR

```{r study population}
# pop contains studypop information (exposed unexposed) along with scrcr variables for all exposed
# studypop contains lopnr, diagdate_scrcr, strata, case
pop <- left_merge(studypop, scrcr_first, by = c("lopnr"))
```

### Comparators

General population comparators. No information from SCRCR is available for comparators.

```{r}
pop[, .N, by = "case"]
```

# 3. Variables

### Start of follow up

`index_date` = `procedure_date` from scrcr

```{r index date}
pop[, index_date := proceduredate]
qplot(pop$index_date)

# add index date to surgeries
surgeries <- left_merge(surgeries_raw, pop[, .(lopnr, index_date)], id_vars = "lopnr")
ileus <- left_merge(ileus_raw, pop[, .(lopnr, index_date)], id_vars = "lopnr")
abd_hernia <- left_merge(abd_hernia_raw, pop[, .(lopnr, index_date)])
```


### Cencoring factors

1. #### Death date

Added from CDR. Used for cencoring.

```{r add CDR}
pop <- left_merge(pop, cdr_raw[, .(lopnr, deathdate)], id_vars = "lopnr")

dead_before_indexdate <- truethy(pop$deathdate < pop$index_date)

if (any(dead_before_indexdate)) {
    warning(glue("deaths before index date should not be present (add CRD)
    Removing {sum(dead_before_indexdate)} rows.
    "))
    pop <- pop[!dead_before_indexdate]
}
```

2. #### Migrations

Added from FoB. Used for cencoring. Only emmigration after `index_date` are of interest

```{r add migration}
migrations <- merge(migrations_raw, pop[, .(lopnr, index_date)], by = "lopnr", all.x = TRUE)
migrations_after_first <- migrations[index_date < migrationdate, ]
setorder(migrations_after_first, lopnr, migrationdate)
migrations_after_first <- migrations_after_first[!duplicated(lopnr)]

# Only care for people that migrate after index_date
emmigrations_after_first <- migrations_after_first[migration_type == "emmigration", .(lopnr, index_date, migrationdate)]

pop <- left_merge(pop, emmigrations_after_first, by = c("lopnr", "index_date"))
```

3. #### Recurrent disease (relapse)

Recurrent disease is assumed to have arissen 6 months prior date given in SCRCR
(relapse=1 from CRC-base, date variable; first of `relapse_loc_date` or `relapse_mets_date` from CRC-Base)

_Also identified from the patient register_

```{r recurrent disease}
first_relapse_date_scrcr <- scrcr_processed %>%
    group_by(lopnr) %>%
    summarise(
        relapse_scrcr_date = min(relapse_loc_date, relapse_mets_date, na.rm = TRUE)
    )

first_relapse_date_patreg <- relapse_raw[relapse_opr == 1 | relapse_ipr == 1, .(
    relapse_patreg = any(truethy(relapse_opr | relapse_ipr)),
    relapse_patreg_date = min(indate, na.rm = TRUE)
),
by = "lopnr"
]

# add index date
first_relapse_date_patreg <- left_merge(first_relapse_date_patreg, pop[, .(lopnr, index_date)], by = "lopnr")


# combine relapses from the two sources
first_relapse <- left_merge(first_relapse_date_patreg, first_relapse_date_scrcr)
first_relapse <- first_relapse %>%
    rowwise() %>%
    mutate(relapse_date = min(relapse_patreg_date, relapse_scrcr_date, na.rm = TRUE))
setDT(first_relapse)


pop <- left_merge(pop, first_relapse[, .(lopnr, index_date, relapse_date)], by = c("lopnr", "index_date"))

# Make sure there are no relapses before procedure date
relapse_before_index <- truethy(pop$relapse_date < pop$index_date)
if (any(relapse_before_index)) {
    print(glue("Removing {sum(relapse_before_index)} rows with relapse before index date (proceduredate) when adding relapses from patreg"))
    pop <- pop[!relapse_before_index]
}
```

### Baseline characteristics

#### SCRCR only
_called exposure variables in the document_
All variables are assessed as prevalent at baseline

 - Body mass index (BMI) (weight/height2  from CRC-base, kg/m2) *make categories*
 - Neoadjuvant radiotherapy (neoadj_rt=1 from CRC-base)
 - Minimal invasive surgery (a2_lapa=1 or a2_robotass=1 from SCRCR) _to be exposure instead?_
 - Converted to open (a2_konv=1 from SCRCR)
 - Stoma formation at resection (a2_skydd=1 or a2_perm=1 from SCRCR)
 - Intraoperative blood-loss (a2_blodn from SCRCR) __perop_bleed = a2_blodn__
 - Pathological tumor stage (tstage_path=1-13 from CRC-base)
 - Location of tumour; right (a2_tumlage=1-5) left (a2_tumlage=6-8) rectum (a2_tumlok_beskrivning=2)
 - Colorectal cancer surgery type (a2_optyp2=01-09 [surgery_type has been derived] from SCRCR)
 - Year of surgery (`proceduredate`)

```{r baseline scrcr, include=TRUE}
# BMI - already in data, make categories
pop[, bmi_cat := case_when(
    bmi < 25 ~ "<25",
    bmi < 30 ~ "25-30",
    bmi >= 30 ~ ">=30",
    TRUE ~ ""
)]

# neoadjuvant radiotherapy
pop[, bs_neoadj_radio := truethy(neoadj_rt)]

# minimal invasive surgery
pop[, bs_minimal_invasive_surgery := case_when(
    truethy(a2_lapa) | truethy(a2_robotass) ~ TRUE,
    TRUE ~ FALSE
)]
pop[, .(n_minimal_invasive_surgery = sum(bs_minimal_invasive_surgery)), by = .(a2_lapa, a2_robotass)]

# Converted to open
pop[, bs_convert_to_open := truethy(a2_konv)]

# stoma formation at resection
pop[, bs_stoma_resection := case_when(
    truethy(a2_skydd) | truethy(a2_perm) ~ TRUE,
    TRUE ~ FALSE
)]
pop[, .(n_stoma = sum(bs_stoma_resection)), by = .(a2_skydd, a2_perm)]

# Intraop blood loss
pop[, bs_perop_bleed := perop_bleed]

# patological tumour stage
pop[, bs_path_t_stage := pt]
pop[, bs_path_t_stage_cat := pt_cat]

# Location of tumour
pop[, bs_location_of_tumour := case_when(
    location == "rectum" ~ "rectum",
    col_location == "right" ~ "right",
    col_location == "left" ~ "left",
    TRUE ~ ""
)]

# Colorectal cancer surgery type (surgery_type)
pop$bs_surgery_type <- pop$surgery_type

# Surgery year
pop[, index_year := year(index_date)]
pop$index_year %>% tabyl()
```


#### Whole population

- Abdominal surgery from 1997 to before the colorectal cancer surgery (op= se list, codes from IPR) before diagdate_scrcr _all that starts with J_
- SBO diagnosis (dia1-dia30=K56.5, K56.6, K56.7 from IPR) from 1997 to 3 months within CRC diagnosis (a1_diagnosdatum from SCRCR)  _added a washout period here of 0.5 months to avoid_
- Abdominal pain: First hospitalisation for abdominal pain (dia1-dia30=R10.0, R10.3, R10.4 from IPR) from 1997 to 3 months before CRC operation (diagdate_scrcr from CRC-Base).
- Education (from LISA)
- Charlson (CCI)

```{r baseline all abdominal surgeries before procedure date}
# Adominal surgeries before procedure date
abd_surgeries_before <- previous_abd_surgeries_raw[indate < diagdate_scrcr & surgerycodes %like% "^J| J"]

abd_surgeries_before$bs_previous_abd_surgery <- 1
pop <- left_merge(pop, abd_surgeries_before, by = c("lopnr", "diagdate_scrcr"))
```

```{r baseline all sbo_3m_before}
# sbo 3 months before crcdiagnosis (ileus dataset)
ileus_3m_before <- ileus[indate < diagdate_scrcr & ileus == 1]
ileus_3m_before <- ileus_3m_before[
    ,
    months_before_diagdate := time_length(as_date(diagdate_scrcr) - as_date(indate), unit = "months")
][
    , .(lopnr, index_date, indate, diagdate_scrcr, ileus, months_before_diagdate)
]
ileus_3m_before <- ileus_3m_before[months_before_diagdate <= 3]
first_ileus_3m_before_diagdate <- ileus_3m_before[, .(
    date_ileus_within_3m = min(indate), bs_previous_sbo_3m_before_diagdate = 1
),
by = c("lopnr", "index_date")
]

# ileus distribution
qplot(first_ileus_3m_before_diagdate$date_ileus_within_3m) + theme_minimal()

pop <- left_merge(
    pop,
    first_ileus_3m_before_diagdate[, .(lopnr, index_date, bs_previous_sbo_3m_before_diagdate)],
    id_vars = c("lopnr", "index_date")
)
```

```{r abdominal pain within 3m before diagdate}
# COntains both hernia and pain
abpain_3m_before <- abd_hernia[abpain == 1]

# Create months before diagdate and select variables
abpain_3m_before <- abpain_3m_before[
    ,
    months_before_diagdate := time_length(as_date(diagdate_scrcr) - as_date(indate), unit = "months")
][
    , .(lopnr, index_date, indate, diagdate_scrcr, abpain, months_before_diagdate)
]
# Within 3 months
abpain_3m_before <- abpain_3m_before[months_before_diagdate <= 3]

first_abpain_3m_before_diagdate <- abpain_3m_before[, .(
    date_abpain_within_3m = min(indate),
    bs_previous_abpain_3m_before_diagdate = 1
), by = c("lopnr", "index_date")]

# abpain distribution
qplot(first_abpain_3m_before_diagdate$date_abpain_within_3m) + theme_minimal()

pop <- left_join(
    pop,
    first_abpain_3m_before_diagdate[, .(lopnr, index_date, bs_previous_abpain_3m_before_diagdate)],
    id_vars = c("lopnr", "index_date")
)
```

```{r education}
most_recent_education <- get_education_from_lisa(lisa_raw, pop)
pop <- left_merge(pop, most_recent_education[, .(lopnr, education_cat)])
```

```{r cci}
cci <- rbindlist(list(
    cci_exposed_raw[, .(lopnr, diagdate_scrcr, CCI_weighted)],
    cci_comparators_raw[, .(lopnr, diagdate_scrcr, CCI_weighted)]
))
colnames(cci) <- str_to_lower(colnames(cci))

pop <- left_merge(pop, cci, by = c("lopnr", "diagdate_scrcr"))
```


#### Occurs after exposure
_Not assessed_

- Re-operation (a4_reop from SCRCR)
- Fascial dehiscence (a4_reoprup from SCRCR)
- Anastomotic leak (a4_anakomp from SCRCR)


### Other SBO mechanisms

#### a. Hernia
Incarcerated abdominal hernia (dia1-dia30= K40.0, K40.3, K41.0, K41.3, K42.0, K43.0, K43.3, K43.6, K44.0, K45.0, K46.0 as primary or secondary codes in the IPR) _at the same visit as SBO_

```{r hernia}
hernia <- abd_hernia[abhernia == 1 & indate < index_date]
hernia_first <- hernia[,
    .(hernia_at_sbo = any(abhernia), hernia_date = min(indate)),
    by = c("lopnr", "index_date")
]

hernia_first$hernia_date %>% qplot()

# Add to population
pop <- left_merge(pop, hernia_first[, .(lopnr, index_date, hernia_at_sbo)], id_vars = c("lopnr", "index_date"))
```


#### b. Stoma closures
 _Moved it to here instead of on its own so assessed at SBO just like hernia, ie at the same visit as SBO. Should it be assessed at baseline? Between exposure and outcome? is it an outcome, a secondary outcome??_

Any procedure code (op=JFG00, JFG10, JFG20, JFG23, JFG29, JFG26,JFG30, JFG33, JFG36 in the IPR)

```{r stoma closures}
# Codes used

definitions$procedure_codes$stoma_closure

# Before and after index_date
stoma_closures <- surgeries[code %in% definitions$procedure_codes[["stoma_closure"]], .(
    lopnr, index_date, indate, code
)]

tabyl(stoma_closures, code)

# TODO : dont know when to assess
```


### Outcomes


#### 1. SBO (small bowel obstruction) after colorectal resection (proceduredate)
- `sbo_scrcr` u5_ileus_v_rde=1 (u0_besoksdatum for date sbo)
- `sbo_patreg` dia k56.5-7 (indatum for date sbo) found in uk_ileus dataset

```{r outcome sbo}

# sbo_scrcr (will be missing for all comparators)
sbo_scrcr <- scrcr_processed[u5_ileus == 1, .(
    sbo_scrcr = any(truethy(u5_ileus)),
    sbo_scrcr_date = min(u0_besoksdat)
), by = "lopnr"]


pop <- left_merge(pop, sbo_scrcr, id_vars = "lopnr")
if (any(pop$index_date >= pop$sbo_scrcr_date)) {
    warning("sbo_scrcr_date before index_date at sbo outcome. This should not happen as only the first scrcr should be included. Need to investigate")
}

# sbo_patreg

# Only keep ileus (also contains hernia and abpain icds)
# All codes are before diagdate
ileus_only <- ileus[ileus == 1]
# only after index_date
ileus_after <- ileus_only[index_date < indate]

# first one sbo from patreg
ileus_first <- ileus_after[, .(
    sbo_patreg = any(ileus == 1),
    sbo_patreg_date = min(indate, na.rm = TRUE)
), by = c("lopnr", "index_date")]

pop <- left_merge(pop, ileus_first, id_vars = c("lopnr", "index_date"))
```

```{r sbo summary}
pop[, .(sbo_scrcr = n_per(sbo_scrcr), sbo_patreg = n_per(sbo_patreg)), by = case]
```


#### 2. SBO-surgery 

__Assuming it does not require a prios SBO diagnosis__

- u5_ileus_reop_v_rde eq 1 (u0_besoksdatum for date sbo surgery)
- alt1 k56.5-7 plus op=JAP00, JAP01, JFK00, JFK01, JFK10, JFK20, JFK96, JFK97 __Only do alt1 for now__
- alt2 k56.5-7 plus op= JAP00, JAP01, JFB00, JFB01, JFB10, JFB13, JFB20, JFB21, 
                        JFB96, JFB97, JFC00, JFC01, JFC10, JFC11, JFC20, JFC21,
                        JFF10, JFF11, JFF13, JFK00, JFK01, JFK10, JFK20, JFK96, JFK97 

```{r outcome sbo-surgery}

# SBO surgery scrcr
sbo_surgery_scrcr <- scrcr_processed[u5_ileus_reop == 1,
    .(
        sbo_surgery_scrcr = any(truethy(u5_ileus == 1)),
        sbo_surgery_scrcr_date = min(u0_besoksdat, na.rm = TRUE)
    ),
    by = c("lopnr")
]
pop <- left_merge(pop, sbo_surgery_scrcr, id_vars = "lopnr")
```

```{r sbo surgeries1}

# SBO sugery patreg --

# only relevant codes after index_date
sbo_surgeries1 <- surgeries %>%
    filter(index_date < indate) %>%
    filter(code %in% definitions$procedure_codes[["sbo_operation_alt1"]]) %>%
    select(lopnr, index_date, indate, hdia:dia30, code)

# Find rows with ileus icd code
ileus_pattern <- definitions$icd_codes$ileus %>% str_c(collapse = "|")

# looks through all the dia columns and returns a vector with one value per row indicating if any ileus icds are present
sbo_surgeries1$ileus_dia <- apply(sbo_surgeries1, 1, function(x) any(str_detect(x, ileus_pattern))) %>% truethy()
sbo_surgeries1 <- sbo_surgeries1[ileus_dia == 1]

# ileus icd combined  with sugery
sbo_surgeries_first1 <- sbo_surgeries1[,
    .(
        sbo_surgery_patreg1 = TRUE,
        sbo_surgery_patreg_date1 = min(indate)
    ),
    by = c("lopnr", "index_date")
]

pop <- left_merge(pop, sbo_surgeries_first1, id_vars = c("lopnr", "index_date"))
```

```{r sbo surgeries2}

# SBO sugery patreg --

# only relevant codes after index_date
sbo_surgeries2 <- surgeries %>%
    filter(index_date < indate) %>%
    filter(code %in% definitions$procedure_codes[["sbo_operation_alt2"]]) %>%
    select(lopnr, index_date, indate, hdia:dia30, code)

# Find rows with ileus icd code
ileus_pattern <- definitions$icd_codes$ileus %>% str_c(collapse = "|")

# looks through all the dia columns and returns a vector with one value per row indicating if any ileus icds are present
sbo_surgeries2$ileus_dia <- apply(sbo_surgeries2, 1, function(x) any(str_detect(x, ileus_pattern))) %>% truethy()
sbo_surgeries2 <- sbo_surgeries2[ileus_dia == 1]

# ileus icd combined  with sugery
sbo_surgeries_first2 <- sbo_surgeries2[,
    .(
        sbo_surgery_patreg2 = TRUE,
        sbo_surgery_patreg_date2 = min(indate)
    ),
    by = c("lopnr", "index_date")
]

pop <- left_merge(pop, sbo_surgeries_first2, id_vars = c("lopnr", "index_date"))
```

```{r sbo_surgery summary}

pop[, .(
    sbo_surgery_scrcr = n_per(sbo_surgery_scrcr),
    sbo_surgery_patreg1 = n_per(sbo_surgery_patreg1),
    sbo_surgery_patreg2 = n_per(sbo_surgery_patreg2)
), by = case]
```


### Secondary outcomes 
Is this right? Are these secondary outcomes? Should 

#### 1. Abdominal pain

First hospitalisation for abdominal pain (dia1-dia30=R10.0, R10.3, R10.4 from IPR) after CRC operation (proceduredate from CRC-Base). 

```{r outcome abdominal pain}
abpain_after <- ileus[abpain == 1 & indate > index_date]

abpain_after[, .(abpain = any(abpain), abpain_date = min(indate)), by = c("lopnr", "index_date")]
```


### End of follow up (eof)

SCRCR outcome variables will only be assessed for exposed (`sbo_scrcr_eof`). 

__Create__:
- `sbo_scrcr_eof`
- `sbo_eof`
- `sbo_surgery_eof1`
- `sbo_surgery_eof2`

Which ever comes first of outcome_date, emmigration or death or Date of cancer recurrence(`relapse_date`) OR end of study (assumin `r params$end_of_study`)

```{r eof}

pop <- pop %>%
    rowwise() %>%
    mutate(
        sbo_scrcr_eof = min(deathdate, migrationdate, sbo_scrcr_date, relapse_date, na.rm = TRUE),
        sbo_surgery_scrcr_eof = min(deathdate, migrationdate, sbo_surgery_scrcr_date, relapse_date, na.rm = TRUE),
        sbo_eof = min(deathdate, migrationdate, sbo_patreg_date, relapse_date, na.rm = TRUE),
        sbo_surgery_eof1 = min(deathdate, migrationdate, sbo_surgery_patreg_date1, relapse_date, na.rm = TRUE),
        sbo_surgery_eof2 = min(deathdate, migrationdate, sbo_surgery_patreg_date2, relapse_date, na.rm = TRUE)
    )
setDT(pop)
```

### Follow up time (fu) 
ie `time_since_proceduredate`
`index_date` to `end of follow up` in months.

```{r fu}

pop[, `:=`(
    sbo_scrcr_fu_m = time_length(as_date(sbo_scrcr_eof) - as_date(index_date), "months"),
    sbo_surgery_scrcr_fu_m = time_length(as_date(sbo_surgery_scrcr_eof) - as_date(index_date), "months"),
    sbo_fu_m = time_length(as_date(sbo_eof) - as_date(index_date), "months"),
    sbo_surgery_fu_m1 = time_length(as_date(sbo_surgery_eof1) - as_date(index_date), "months"),
    sbo_surgery_fu_m2 = time_length(as_date(sbo_surgery_eof2) - as_date(index_date), "months")
)]
```

```{r sbo fu summary}
plot_ly(data = pop) %>%
    add_trace(x = ~sbo_scrcr_fu_m, type = "histogram", name = "SBO scrcr") %>%
    add_trace(x = ~sbo_fu_m, type = "histogram", name = "SBO patreg")
```

```{r sbo surgery fu summary}
plot_ly(data = pop) %>%
    add_trace(x = ~sbo_surgery_scrcr_fu_m, type = "histogram", name = "SBO surgery scrcr") %>%
    add_trace(x = ~sbo_surgery_fu_m1, type = "histogram", name = "SBO surgery1") %>%
    add_trace(x = ~sbo_surgery_fu_m2, type = "histogram", name = "SBO surgery2")
```

### Follow up time categorized 

ie `time since proceduredate`
I dont get why start at 6 months here? _added <6m_

Three time periods (6-18 months, 18-41 months, 41-66 months)

__Done for__: 

- `sbo_fu_m` into `sbo_fu_cat_m`
- `sbo_surgery_fu_m` into `sbo_surgery_fu_cat_m`

```{r time_since_procedure}

pop[, `:=`(
    sbo_scrcr_fu_cat_m = case_when(
        sbo_scrcr_fu_m < 0 ~ "1. <0m",
        sbo_scrcr_fu_m < 18 ~ "2. 6m < 18m",
        sbo_scrcr_fu_m < 41 ~ "3. 18m < 41m",
        sbo_scrcr_fu_m < 66 ~ "4. 41m < 66m",
        sbo_scrcr_fu_m >= 66 ~ "5. >=66m",
    ),
    sbo_fu_cat_m = case_when(
        sbo_fu_m < 0 ~ "1. <0m",
        sbo_fu_m < 18 ~ "2. 6m < 18m",
        sbo_fu_m < 41 ~ "3. 18m < 41m",
        sbo_fu_m < 66 ~ "4. 41m < 66m",
        sbo_fu_m >= 66 ~ "5. >=66m",
    ),
    sbo_surgery_fu_cat_m1 = case_when(
        sbo_surgery_fu_m1 < 0 ~ "1. <0m",
        sbo_surgery_fu_m1 < 18 ~ "2. 6m < 18m",
        sbo_surgery_fu_m1 < 41 ~ "3. 18m < 41m",
        sbo_surgery_fu_m1 < 66 ~ "4. 41m < 66m",
        sbo_surgery_fu_m1 >= 66 ~ "5. >= 66m"
    ),
    sbo_surgery_fu_cat_m2 = case_when(
        sbo_surgery_fu_m2 < 0 ~ "1. <0m",
        sbo_surgery_fu_m2 < 18 ~ "2. 6m < 18m",
        sbo_surgery_fu_m2 < 41 ~ "3. 18m < 41m",
        sbo_surgery_fu_m2 < 66 ~ "4. 41m < 66m",
        sbo_surgery_fu_m2 >= 66 ~ "5. >= 66m"
    ),
    sbo_surgery_scrcr_fu_cat_m = case_when(
        sbo_surgery_scrcr_fu_m < 0 ~ "1. <0m",
        sbo_surgery_scrcr_fu_m < 18 ~ "2. 6m < 18m",
        sbo_surgery_scrcr_fu_m < 41 ~ "3. 18m < 41m",
        sbo_surgery_scrcr_fu_m < 66 ~ "4. 41m < 66m",
        sbo_surgery_scrcr_fu_m >= 66 ~ "5. >= 66m"
    )
)]
tabyl(pop, sbo_fu_cat_m) %>% arrange(1)
tabyl(pop, sbo_surgery_fu_cat_m1) %>% arrange(1)
```

## Export analysisdata

```{r export}
print(glue("Writing analysisdata with : {nrow(pop)} rows and {ncol(pop)} columns to: project derived folder"))
write_to_project_derived(pop, "analysisdata.csv")
```